\chapter{Bilingual Language Profile: Nanjing Dialect - Mandarin}	% *NOT* \OnePageChapter
\label{appendix:BLP}

What follows is an English translation of the text shown to participants as they filled out this experiment's version of the Bilingual Lanuage Profile \citep{birdsong2012bilingual,gertken2014assessing}.

\section{Instructions}
	We would like to ask you to help us by answering the following questions concerning your language history, use, attitudes, and proficiency. This document is based on a survey created with support from the Center for Open Educational Resources and Language Learning at the University of Texas at Austin. Its purpose is to better understand the profiles of bilingual speakers in diverse settings with diverse backgrounds. The document consists of 19 questions and will take less than 10 minutes to complete. This is not a test, so there are no right or wrong answers. Please answer every question and give your answers sincerely.
	
	Thank you very much for your help.
	
	Please click "NEXT" to begin filling out the form.

\section{Biographical Information}
Subject ID: $\rule{5cm}{0.15mm}$

Today's Date $\rule{1cm}{0.15mm}$ / $\rule{1cm}{0.15mm}$ / $\rule{1cm}{0.15mm}$

Age $\rule{1cm}{0.15mm}$

Male / Female / Prefer not to say / Other: $\rule{3cm}{0.15mm}$

Nanjing Dialect District $\rule{5cm}{0.15mm}$

Highest level of formal education:

\indent \indent Primary or middle school

\indent \indent High school

\indent \indent Some college

\indent \indent College (B.A., B.S.)

\indent \indent Some graduate school

\indent \indent Master's

\indent \indent PhD / MD / JD

\indent \indent Other: $\rule{3cm}{0.15mm}$