\chapter{Discussion}
\label{discussionchapter}

It seems to me that I don't have good evidence for a direct, mechanistic link between interpersonal accommodation and real-world sound change, at least not one like Trudgill suggests. In my opinion, the fact that different variables behave differently in the experiment argues against that model, at least the model that says ``the more numerous speakers will just overwhelm the less numerous ones.'' I don't have solid numbers about how numerous the speakers of the dialects are across the city, but consider that:

\begin{enumerate}
    \item there's a Mandarin dominance bias in my subject pool, I'm pretty sure both in terms of the BLP results \textit{and} the IAT results;
    \item anecdotally, people who don't speak any \ND{} in the city WAAAAY out number people who don't speak any Mandarin.
\end{enumerate}

I think ``mechanistic'' is way too strong a word. ``Probabilistic'' is almost certainly better, but i don't really have a strong reason to argue for that particular term.

It's true that the way my experiment works is not exactly the same way that accommodation works in the real world, but it seems unlikely to me that the relationship between lab accommodation and real-world sound change should be the exact opposite of real-world accommodation and real-world sound change.

I recently realized that the only reason you don't see lots of studies of sound change from the lab that \emph{really} get at the link between the micro and the macro is that first you do some manipulation to an individual and watch how they behave, or you follow the changes of a larger group, but without any particularly solid manipulation. What if you were to figure out how a social network were structured, then primed specific people in the network with, say, extended VOT, and then let the network chatter and watch how the change behaves. I feel like I've seen a computational version of basically this where bots start of with one phonological vowel, but vary and sometimes imitate, and after like 80\% of simulations they ended up with a 5 vowel system.

\pagebreak

This page intentionally left blank.